A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-24 text/manuscript.Rmd
add acknowledgements, cut survival section a little more

To discover what pathways drive the UMAP clustering, we calculated percent
enrichment for each pathway _within cluster_ relative to all other clusters
combined (see methods) and ranked pathways for each cluster from highest to
lowest enrichment. We visualized the enrichment as a heatmap (**figure 5B**).
Using this approach, we identified 25 pathways highly enriched (enrichment score
$> 0.3$, see methods) in cluster 2, 9 pathways enriched in cluster 3, 24
pathways in cluster 4, 10 pathways in cluster 5, 10 pathways in cluster 6, 24
pathways in cluster 7, 24 pathways in cluster 8, 10 pathways in cluster 9, and 0
pathways for clusters 1 and 10. Clusters 7, 8, and 9 had several pathways in
common. To explore the specific pathways marking each cluster, we projected
disruptions for each of the 377 pathways onto the UMAP (**supplementary figure
S4**). Clusters 3 and 5 were distinguished by metabolic pathways including RNA
and protein biosynthesis (**supplementary figure S4**). Similarly, cluster 4 was
distinguished by mutations affecting regulation of DNA and histone methylation
("DNA methylation", "PRC2 methylates histones and DNA", and "Nucleosome
assembly"). Clusters 7-9 have in common mutations in extracellular,
intracellular, and immune-related signaling pathways (see **figure 5C** and
**supplementary figure S4**). Cluster 2 had the highest pathway enrichment
levels of the three, having mutations in hedgehog signaling, "β-catenin
degradation", "cellular response to hypoxia", "regulation of cell cycle" and
"apoptosis" among others.

@@

We used Bayesian inference to explore the relative impact of cancer diagnosis on
survival at the time of diagnosis. Our model assumes that at birth each person
has a small initial probability per unit time to die, and that this probability
grows exponentially in time at some baseline rate that matches the observed CDC
data. The rate of growth in risk is assumed to jump to a higher level at
diagnosis, which we take as a proxy for onset. This represents an accelerated
aging type model. To fit the model, we constrain the baseline risks using the
available CDC life tables. We then constrain the cancer specific multiplier
using the longevity data within the cancer tumor dataset. Our model estimates
the effect of cancer type and class-specific cancer effects independently,
resulting in a cancer and class specific estimate of the aging rate multiplier.
A detailed description of the CDC longevity model and likelihood is provided in
the supplementary methods.

@@

@@

The authors have nothing to declare. This work was supported by a grant from the Cedars-Sinai Precision Health Initiative.

>>>>

To discover what pathways are most important for clustering, we calculated
percent enrichment for each pathway _within cluster_ relative to all other
clusters combined (see methods) and ranked pathways for each cluster from
highest to lowest enrichment. We visualized the enrichment as a heatmap
(**figure 5B**). Using this approach, we identified 25 pathways highly enriched
(enrichment score $> 0.3$, see methods) in cluster 2, 9 pathways enriched in
cluster 3, 24 pathways in cluster 4, 10 pathways in cluster 5, 10 pathways in
cluster 6, 24 pathways in cluster 7, 24 pathways in cluster 8, 10 pathways in
cluster 9, and 0 pathways for clusters 1 and 10. Clusters 7, 8, and 9 had
several pathways in common. To explore the specific pathways marking each
cluster, we projected disruptions for each of the 377 pathways onto the UMAP
(**supplementary figure S4**). Clusters 3 and 5 were distinguished by metabolic
pathways including RNA and protein biosynthesis (**supplementary figure S4**).
Similarly, cluster 4 was distinguished by mutations affecting regulation of DNA
and histone methylation ("DNA methylation", "PRC2 methylates histones and DNA",
and "Nucleosome assembly"). Clusters 7-9 have in common mutations in
extracellular, intracellular, and immune-related signaling pathways (see
**figure 5C** and **supplementary figure S4**). Cluster 2 had the highest
pathway enrichment levels of the three, having mutations in hedgehog signaling,
"β-catenin degradation", "cellular response to hypoxia", "regulation of cell
cycle" and "apoptosis" among others.

@@

We used Bayesian inference to explore the relative impact of cancer diagnosis on
survival at the time of diagnosis. Our model assumes that at birth each person
has a small initial probability per unit time to die, and that this probability
grows exponentially in time at some baseline rate that matches the observed CDC
data. The rate of growth in risk is assumed to jump to a higher level at
diagnosis, which we take as a proxy for onset. This represents an accelerated
aging type model. To fit the model, we constrain the baseline risks using the
available CDC life tables. We then constrain the cancer specific multiplier
using the longevity data within the cancer tumor dataset. Our model estimates
the effect of cancer type and class-specific cancer effects independently,
resulting in a cancer and class specific estimate of the aging rate multiplier.
A detailed description of the CDC longevity model and likelihood is provided in
the **Supplementary methods**.

@@

Wish wish to thank David Van Valen, Simon Knott and Kate Lawrenson for critical
reading of this manuscript, and Ivetth Corona for early discussions and
feedback.

@@

The authors have no conflicts of interest to declare. This work was supported by
a grant from the Cedars-Sinai Precision Health Initiative.
